Non-analogues in paleoecological reconstruction

### 2015-07-30 Paper_Markdown.Rmd
Minor changes from Sakari's comments on the PDF.  Added some methods papers.  Will add more.

To analyse the effects of non-analogue pollen on model skill we chose five pollen-based reconstruction methods from the literature and apply them to the modern pollen data from eastern North America.  The models are MAT, WA, WAPLS, BRT and randomForest.  Each of these models has been used previously in the pollen literature, and has strengths and weaknesses.  MAT [@overpeck1985quantitative] is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.  WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer from a basic assumption of unimodal taxon distributions along climate gradients.

@@

SAKARI or MISKA can you add text here about BRT?

@@

Mean minimum dissimilarity for pollen samples in the modern pollen dataset ranges from `r min(min.diss)` to `r min(min.diss)`, with a 95% confidence interval from `r quantile(min.diss, 0.025)` to `r quantile(min.diss, 0.975)`.  The highest minimum analogue distance is for an unpublished site along the New York, Pennsylvannia border.  The site is notable for very high proportions of *Fraxinus pennsylvanica/americana* pollen, more than 5x higher than the next closest site.  The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al [-@williams2009rapid] and XXXX.  The geographic distribution of non-analogue sites is particularly interesting since many of the sites are in close proximity to other modern sites.  This indicates that While spatial autocorreltaion is likely to play a role in the proximity of analogues (and thus their predictions) [@telford2009evaluation] it is not always the case that spatially proximate samples are compositionally proximate, and indeed, pollen assemblages in densely sampled regions may exhibit no-analogue characteristics.

>>>>

To analyse the effects of non-analogue pollen on model skill we chose five pollen-based reconstruction methods from the literature and apply them to the modern pollen data from eastern North America.  The models are MAT [@overpeck1985quantitative], WA [@ter1993weighted], WAPLS [@ter1989inferring], BRT and randomForest.  Each of these models has been used previously in the pollen literature, and has strengths and weaknesses.  MAT is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.  WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer from a basic assumption of unimodal taxon distributions along climate gradients.

@@

BRT [@de2007boosted] and randomForest [@breiman2001random] are ensemble machine-learning approaches which have seen increasing use in fossil pollen literature during the past few years [][SOME REFS HERE IF NECESSARY ... GORING ET AL. 2010, ZHAO ET AL. 2012, SALONEN ET AL. 2014...]. Both methods are based on creating an ensemble of regression tree models with all pollen taxa used as predictors and an environmental variable of choice as the response. The randomForest creates an ensemble of independent tree models (here 500 trees), while using bootstrap samples of the calibration data when building each tree, and randomly selected subsets of the predictor set when determining each tree split, thus creating diversity in the tree ensemble. In BRT, by comparison, the tree ensemble is built sequentially, with each added tree aiming to explain the residuals of the previously fitted ensemble. The final BRT model can thus be likened to an additive regression model in which the individual terms are regression trees.

@@

Mean minimum dissimilarity for pollen samples in the modern pollen dataset ranges from `r min(min.diss)` to `r max(min.diss)`, with a 95% confidence interval from `r quantile(min.diss, 0.025)` to `r quantile(min.diss, 0.975)`.  The highest minimum analogue distance is for an unpublished site along the New York, Pennsylvannia border.  The site is notable for very high proportions of *Fraxinus pennsylvanica/americana* pollen, more than 5x higher than the next closest site.  The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al [-@williams2009rapid] and XXXX.  The geographic distribution of non-analogue sites is particularly interesting since many of the sites are in close proximity to other modern sites.  This indicates that While spatial autocorrelation is likely to play a role in the proximity of analogues (and thus their predictions) [@telford2009evaluation] it is not always the case that spatially proximate samples are compositionally proximate, and indeed, pollen assemblages in densely sampled regions may exhibit no-analogue characteristics.
